---
title: "Testing Discovery Rate"
author: "Eric Reed"
date: "1/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

# Requirements

## R Packages and Python Modules

```{r}

## R packages to run simulation functions
require(magrittr)
library(igraph)
library(reticulate)
library(Matrix)
library(BDgraph)
library(ggplot2)
require(corpcor)
require(MASS)
require(Biobase)
require(ConAn)

## R packages for visualization
require(ggplot2)
require(gridExtra)

## Python modules
nx <- import("networkx")

```

## Directories

```{r}

baseDir <- ".." # If running from the "./docs" directory
rDir <- file.path(baseDir, "R")
pyDir <- file.path(baseDir, "python")

```


## Read in functions

```{r}

## Read in R functions
rDir %>%
    list.files(., full.names = TRUE) %>%
    lapply(., source) %>%
    invisible # This keeps lapply from printing NULL

## Read in python functions
source_python(file.path(pyDir, "LFR_Communities.py"))

```

# Discovery rate for a single set of parameters

## Single set of parameters

```{r}

# Seed
seed <- 123

# Network
nFeats <- 500
minComSize <- 25
maxComSize <- 200
aveDeg <- 5

# Correlation matrix
Dval <- 0.75

# Number of samples per group
n0 <- 50
n1 <- 50

# Number of iterations and cores for running ConAn
nIter = 200
nCores = 2

```

## Simulate network and correlation matrix

```{r}

# Simulate network
netComs <- simNetComs(seed=seed, 
                      n=nFeats, 
                      tau1=3, 
                      tau2=2, 
                      mu=0.07, 
                      average_degree=aveDeg, 
                      min_community=minComSize, 
                      max_community=maxComSize)

## Create D matrix
Dmat <- matrix(Dval, ncol = length(netComs$coms), nrow = length(netComs$coms))
diag(Dmat) <- 1

## Use D matrix in correlation estimates
set.seed(seed)
netConn <- simAdjMat(netComs, D = Dmat)

## Print community and background connectivity
print(netConn$conns)
print(netConn$bg)

```

## Simulate data set of groups from same correlation matrix

```{r}

## Two of the same correlation matrices
cor0 <- cor1 <- netConn$adjMat

## Simulate expression set
eSet <- eSetSim(n0, n1, cor0, cor1)

## Print IDs in data set

### Sample IDs
print(head(colnames(eSet)))

### Feat IDs
print(head(colnames(eSet)))

### Group IDs
print(table(eSet$group))

## Create module list
modList <- lapply(levels(netConn$coms), function(x) rownames(eSet)[netConn$coms == x])
names(modList) <- levels(netConn$coms)

```

## Calculate p-values for each module and ConAn run

```{r}

Pres <- conanWrapperPval(eSet, modList, nIter, nCores, seed)

## Add some relavant information
Pres$conn1 <- Pres$conn0 <- netConn$conns
Pres$bg1 <- Pres$bg0 <- netConn$bg

## Print top of data frame
print(head(Pres))

## Print p-values for module 1
print(Pres[Pres$mod == "1",])

```

## Change background connectivity in one group

```{r}

## Set modular connectivity in the correlation matrix for one group to NA
corComRem <- cor1 # or cor0
for(i in levels(netConn$coms)) corComRem[netConn$coms == i, netConn$coms == i] <- NA

## Get shrink factor to reduce background connectivity to a value of 0.3
shrVal <- findShrinkFactor(0.3, corComRem)

## Shrink correlaton matrix for one group
cor_bgRed <- shrinkCor(cor1, shrVal) # or com0

### Community
getComConn(levels(netConn$coms), cor_bgRed, netConn$coms)

### Background
getComConn("bg", cor_bgRed, netConn$coms)

## Simulate data using this new matrix
eSetBG <- eSetSim(n0, n1, cor0, cor_bgRed)

## Calculate p-values
PresBG <- conanWrapperPval(eSetBG, modList, nIter, nCores, seed)

## Print new connectivity values

## Add some relavant information
PresBG$conn1 <- PresBG$conn0 <- netConn$conns
PresBG$bg0 <- netConn$bg
PresBG$bg1 <- getComConn("bg", cor_bgRed, netConn$coms)

## Print top of data frame
print(head(PresBG))

## Print p-values for module 1
print(PresBG[PresBG$mod == "1",])

```

## Change connectivity of each module in one group

```{r}

## Get targets for each module
targets <- netConn$conns - 0.1

## Get shrinking factors for each module
shrVec <- sapply(levels(netConn$coms), function(x) {
    corModx <- cor1[netConn$coms == x, netConn$coms == x]
    findShrinkFactor(targets[x], corModx)
})

## Shrink each module
cor_modRed <- cor1 # or com0
for(i in levels(netConn$coms)) {
    
    cor_modRed[netConn$coms == i, netConn$coms == i] <- shrinkCor(cor1[netConn$coms == i, netConn$coms == i], shrVec[i])
    
    # This doesn't always work.
    if(!is.positive.definite(cor_modRed)) {
      cor_modRed[netConn$coms == i, netConn$coms == i] <- cor1[netConn$coms == i, netConn$coms == i]
    }
    
}

## Print new connectivity values

### Community
getComConn(levels(netConn$coms), cor_modRed, netConn$coms)

### Community
getComConn("bg", cor_modRed, netConn$coms)

## Simulate data using this new matrix
eSetDiffConn <- eSetSim(n0, n1, cor0, cor_modRed)

## Calculate p-values
PresDiffConn <- conanWrapperPval(eSetDiffConn, modList, nIter, nCores, seed)

## Add some relavant information
PresDiffConn$conn0 <- netConn$conns
PresDiffConn$conn1 <- getComConn(levels(netConn$coms), cor_modRed, netConn$coms)
PresDiffConn$bg1 <- PresDiffConn$bg0 <- netConn$bg

## Remove instances where the connectivity could not be reduced
PresDiffConn <- PresDiffConn[PresDiffConn$conn0 != PresDiffConn$conn1,]

## Print top of data frame
print(head(PresDiffConn))

## Print p-values for module 1
print(PresDiffConn[PresDiffConn$mod == "1",])

```

## Combining differential connectivity and background connectivity

### Situation 1: Smaller background connectivity in group with greater modular connectivity

We already have everything we need:

1. A correlation matrix with reduced background connectivity, *cor_bgRed*.
2. A correlation matrix with reduced modular connectivity, *cor_modRed*.

```{r}

## Simulate data using this new matrix
eSetBG0_MR1 <- eSetSim(n0, n1, cor_bgRed, cor_modRed)

## Calculate p-values
PresBG0_MR1 <- conanWrapperPval(eSetBG0_MR1, modList, nIter, nCores, seed)

## Add some relavant information
PresBG0_MR1$conn0 <- netConn$conns
PresBG0_MR1$conn1 <- getComConn(levels(netConn$coms), cor_modRed, netConn$coms)
PresBG0_MR1$bg0 <- getComConn("bg", cor_bgRed, netConn$coms)
PresBG0_MR1$bg1 <- netConn$bg

## Remove instances where the connectivity could not be reduced
PresBG0_MR1 <- PresBG0_MR1[PresBG0_MR1$conn0 != PresBG0_MR1$conn1,]

## Print top of data frame
print(head(PresBG0_MR1))

## Print p-values for module 1
print(PresBG0_MR1[PresBG0_MR1$mod == "1",])

```

### Situation 2: Smaller background connectivity in group with smaller modular connectivity

Here, we'll just need to reduce background connectivity in the group with reduced modular connectivity.

```{r}

## Shrink correlaton matrix for group
cor_modRed_bgRed <- shrinkCor(cor_modRed, shrVal)

## Simulate data using this new matrix
eSetBG1_MR1 <- eSetSim(n0, n1, cor0, cor_modRed_bgRed)

## Calculate p-values
PresBG1_MR1 <- conanWrapperPval(eSetBG1_MR1, modList, nIter, nCores, seed)

## Add some relavant information
PresBG1_MR1$conn0 <- netConn$conns
PresBG1_MR1$conn1 <- getComConn(levels(netConn$coms), cor_modRed, netConn$coms)
PresBG1_MR1$bg0 <- netConn$bg
PresBG1_MR1$bg1 <- getComConn("bg", cor_modRed_bgRed, netConn$coms)

## Remove instances where the connectivity could not be reduced
PresBG1_MR1 <- PresBG1_MR1[PresBG1_MR1$conn0 != PresBG1_MR1$conn1,]

## Print top of data frame
print(head(PresBG1_MR1))

## Print p-values for module 1
print(PresBG1_MR1[PresBG1_MR1$mod == "1",])

```


# Combine results into a single data frame

```{r}

simRes <- rbind(Pres, PresDiffConn, PresBG, PresBG0_MR1, PresBG1_MR1)

## You can use the seed as simulation ID
simRes$id <- seed

## Check size of output
print(dim(simRes))

```

## Which simulation parameters did we test

```{r}

## Subset for one set of ConAn parameters
simResBFF <- simRes[simRes$sim_type == "bootstrap" & 
                      simRes$mean_correct == FALSE &
                      simRes$mdc_type == "fraction",]

## Add rounded module connectivity
simResBFF$conn05 <- round(netConn$conns[as.character(simResBFF$mod)]/0.05) * 0.05

## Add rounded background connectivity
simResBFF$bg05 <- round(netConn$bg/0.05) * 0.05

## Add rounded differential modular connectivity 
simResBFF$diff05 <- round((simResBFF$conn0 - simResBFF$conn1)/0.05) * 0.05

## Add rounded differential background connectivity 
simResBFF$diffbg05 <- round((simResBFF$bg0 - simResBFF$bg1)/0.05) * 0.05

## Get fequency table
Ftab <- as.data.frame(table(simResBFF$conn05, simResBFF$bg05, simResBFF$diff05, simResBFF$diffbg05))
colnames(Ftab) <- c("conn", "bg", "diff", "diffbg", "count")
Ftab <- Ftab[Ftab$count != 0,]

print(Ftab)
```


